Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs

### 2016-09-06 redaction/slides_LEDa.tex
Nouvelle version encore

\date[September 2016]{ETSG, Helsinki, September 2016}
\title[Trade Costs Black Box]{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}
\author[Daudin et al.]{Guillaume Daudin\\
{\footnotesize Universit\'{e} Paris-Dauphine, PSL \& OFCE }\\ \smallskip
J\'{e}rome H\'{e}ricourt \\
{\footnotesize Universit\'{e} de Lille I \& CEPII }\\  \smallskip
Lise Patureau \\
{\footnotesize  Universit\'{e} Paris-Dauphine, PSL}}

@@

\begin{frame}
\frametitle{Three questions, three answers}
\begin{itemize}
\item What is the size of the iceberg and the additive trade costs? \vspace{0.1cm}
\item[$\Rightarrow$] Provide a quantitative measure of both  \vspace{0.1cm}
\begin{itemize}
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in air and ocean transport (mean value over 1974-2013) \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price  \vspace{0.1cm}
\end{itemize}
\item What do we lose by skipping the additive part of transport costs? \vspace{0.1cm}
\item[$\Rightarrow$] We lose much \vspace{0.1cm}
\begin{itemize}
\item[-] With the additive term included, the ad-valorem component reduces by a factor of 2 \vspace{0.1cm}
\item[-] With the additive term taken into account, a significantly better ``goodness-of-fit'' \vspace{0.1cm}
\end{itemize}
\item How have international transport costs evolved over time? \vspace{0.1cm}
\begin{itemize}
\item[-] Transport costs \textit{per se} have fallen since 1985, by 40\% \vspace{0.1cm}
\item[-] When additive costs are included, not much difference between air and sea transport
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame} [label=slide_data]
\frametitle{Data sources}
\begin{itemize}
%\item Measure international transport costs as the difference between the export price and the import price \vspace{0.1cm}
\item Database: US annual imports of merchandise database \hyperlink{app_data}{\beamergotobutton{More}}\vspace{0.1cm}
\begin{itemize}
\item[-] Yearly basis, from 1974 to 2013, HS 10 digit classification level \vspace{0.1cm}
\item[-] The export (fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export gate \vspace{0.1cm}
\item[-] The import (cif) price, $p$: the price for one kg of merchandise at the US import gate \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Our dependent variable: The ratio $p/\widetilde{p}$ \vspace{0.1cm}
\begin{itemize}
\item[-] At the 3-digit classification level  \vspace{0.1cm}
\item[-] Estimation at the 4-digit level on some selected years as robustness \vspace{0.1cm}
\item[-] Approximatively 200 products (3 digits), from around 200 countries  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] Around 600-700 products at the 4-digit level
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (1)}
\textbf{The estimated equation}
\begin{itemize}
\item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
\item For product $k$, from country $i$ (and by transport mode, air or vessel) \vspace{0.1cm}
\item Rewrite to get:
$$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 (at the $k=$ 3-digit classification level)
\end{itemize}
\end{frame}

@@

\begin{frame} [label=slide_method]
\begin{itemize}
\item Taking logs, we finally estimate the following equation
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
\end{equation}
\normalsize
\item A non-linear equation (due to the additive costs) $\rightarrow$ Estimation using non-linear squares \hyperlink{app_method}{\beamergotobutton{More}} \vspace{0.1cm}
\item How to characterize the importance of additive costs relatively to iceberg? \vspace{0.1cm}
\item[$\Rightarrow$] Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
\begin{itemize}
\item[-] For each year and by transport mode, estimate two equations, depending on additive costs included or not \vspace{0.1cm}
\end{itemize}
\item When additive costs are included:
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 +\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_with_add}
\end{equation}
\normalsize
\begin{itemize}
\item[-] From which we re-built:
$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item When assuming iceberg costs only
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
\end{equation}
\begin{itemize}
\item[-] From which we re-built:
$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
\end{itemize}
\item  Taking the average over the product-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
\begin{itemize}
\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{\tau}^{add}$  \vspace{0.1cm}
\item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
\end{itemize}
\end{itemize}
\end{frame}

>>>>

\date[September 2016]{S\'{e}minaire LEDa, Université Paris Dauphine, Septembre 2016}
\title[Trade Costs Black Box]{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}
\author[Daudin et al.]{Guillaume Daudin\\
{\footnotesize Universit\'{e} Paris-Dauphine, PSL \& OFCE }\\ \smallskip
J\'{e}rome H\'{e}ricourt \\
{\footnotesize Universit\'{e} de Lille I \& CEPII }\\  \smallskip
Lise Patureau \\
{\footnotesize  Universit\'{e} Paris-Dauphine, PSL}}

@@

\begin{frame}
\frametitle{Three questions, three answers}
\begin{itemize}
\item What is the size of the iceberg and the additive trade costs? \vspace{0.1cm}
\item[$\Rightarrow$] Provide a quantitative measure of both  \vspace{0.1cm}
\begin{itemize}
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in air and ocean transport (mean value over 1974-2013) \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price  \vspace{0.1cm}
\end{itemize}
\item What do we lose by skipping the additive part of transport costs? \vspace{0.1cm}
\item[$\Rightarrow$] We lose much \vspace{0.1cm}
\begin{itemize}
\item[-] With the additive term included, the ad-valorem component reduces by a factor of 2 \vspace{0.1cm}
\item[-] With the additive term taken into account, a significantly better ``goodness-of-fit'' \vspace{0.1cm}
\end{itemize}
\item How have international transport costs evolved over time? \vspace{0.1cm}
\begin{itemize}
\item[-] Transport costs \textit{per se} have fallen since 1985, by 40\% \vspace{0.1cm}
\item[-] When additive costs are included, not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2011)
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame} [label=slide_data]
\frametitle{Data sources}
\begin{itemize}
\item Our measure of international transport costs: The difference between the export price and the import price \vspace{0.1cm}
\item Database: US Imports of Merchandise database \hyperlink{app_data}{\beamergotobutton{More}}\vspace{0.1cm}
\begin{itemize}
\item[-] The export (fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export gate \vspace{0.1cm}
\item[-] The import (cif) price, $p$: the price for one kg of merchandise at the entry in the US \vspace{0.1cm}
\item[-] Yearly basis, from 1974 to 2013, HS 10 digit classification level, by transport mode (air or vessel) \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Our dependent variable: The ratio $p/\widetilde{p}$ \vspace{0.1cm}
\begin{itemize}
\item[-] At the 3-digit classification level  \vspace{0.1cm}
\item[-] Estimation at the 4-digit level on some selected years as robustness \vspace{0.1cm}
\item[-] Approximatively 200 products (3 digits), from around 200 countries  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] Around 600-700 products at the 4-digit level
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (1)}
\textbf{The estimated equation}
\begin{itemize}
\item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
\item For product $k$, from country $i$  \vspace{0.1cm}
\item Rewrite to get:
$$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 (at the $k=$ 3-digit classification level) \vspace{0.1cm}
\begin{itemize}
\item[-] Notice the equation is also year- and mode (air or vessel)- specific
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame} [label=slide_method]
\begin{itemize}
\item Taking logs, we finally estimate the following equation
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
\end{equation}
\normalsize
\item A non-linear equation (due to the additive costs) $\rightarrow$ Estimation using non-linear squares \hyperlink{app_method}{\beamergotobutton{More}} \vspace{0.1cm}
\item How to characterize the importance of additive costs relatively to iceberg? \vspace{0.1cm}
\item[$\Rightarrow$] Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
\begin{itemize}
\item[-] For each year and by transport mode, estimate two equations, depending on additive costs included or not \vspace{0.1cm}
\end{itemize}
\item When additive costs are included:
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 +\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_with_add}
\end{equation}

\begin{frame}
\begin{itemize}
\item When assuming iceberg costs only
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
\end{equation}

\item From which we re-built: \vspace{0.1cm}
\begin{itemize}
\item[-] With additive costs:
$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
\item[-] With only iceberg costs:
$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
\end{itemize}
\item  Taking the average over the product-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
\begin{itemize}
\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{\tau}^{add}$  \vspace{0.1cm}
\item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
\end{itemize}
\end{itemize}
\end{frame}
